\thetitle

### 2021-03-03 writing/challenges.tex
touch ups in section 4

While at first glance the method of section~\ref{sec:why-rational} seems straightforward to implement and test,
three challenges remain. The first is comparing the effect of blame on the rational
programmer across three different mechanisms; second is the collection of a large number
of representative debugging scenarios; and third is finding a proper sampling strategy for the
huge space of possibilities. A coincidental challenge is the disparity of the
implementations of gradually typed languages. To eliminate this variable, the
authors use Racket, which is thus far the only language in which all three major
soundness mechanisms are available in a reasonably robust and comparable manner;
for the implementation of the transient semantics, see the parallel submission
of an experience report~\cite{ttt21}.

@@

However, attributing the success of the rational programmer to this or
that blame mechanism demands a careful look at the interplay between blame
and the run-time checks of each gradual type system. When a check fails
the Natural and Transient semantics assign blame instead of using the
information in the exception from the run-time check. Nevertheless, the
exception information may be equally or even more useful to the rational
programmer than blame. If this is the case, then blame may not play a critical
role for the rational programmer. Indeed, precisely because they do not
account for confounding factors, \citet{lksfd-popl-2020} cannot draw any
conclusions about blame specifically despite advertisements for the
opposite. Their experiment can conclude that so-called blame-shifting
works but not that this is because of blame.

Modes offer a uniform way to compare the different semantics and  isolate
blame from the effect of the semantics' runtime checks. Specifically, the Natural and
Transient rational programmer come with a blame mode and an exception
mode. If the former succeeds debugging a program while the latter
fails, then it is safe to conclude that blame is indeed necessary for the
rational programmer. Put differently, the exception mode serves as the
{\em baseline\/} for blame's value within a given semantics; if the
programmer in this mode performs as well or better than the  blame one, a blame assignment
mechanism might be useless.

>>>>

While at first glance the method of section~\ref{sec:why-rational} seems straightforward to implement and test,
three challenges remain. The first is comparing the effect of blame on the rational
programmer across three different mechanisms; second is the collection of a large number
of representative debugging scenarios; and third is finding a proper sampling strategy for the
huge space of possibilities. A coincidental challenge is the disparity of the
implementations of gradually typed languages. To eliminate this variable, the
authors use Racket, which is thus far the only language in which all three major
soundness mechanisms are available in a robust and comparable manner;
for the implementation of Transient Typed Racket, see the parallel submission
of an experience report~\cite{ttt21}.

@@

However, attributing the success of the rational programmer to this or
that blame mechanism demands a careful look at the interplay between blame
and the run-time checks of each gradual type system. When a check fails
the Natural and Transient semantics assign blame instead of using the
information in the exception from the run-time check. Nevertheless, the
exception information may be equally or even more useful to the rational
programmer than blame. If this is the case, then blame may not play a critical
role for the rational programmer. Indeed, precisely because they do not
account for such confounding factors, \citet{lksfd-popl-2020} cannot draw any
conclusions about blame specifically despite advertisements for the
opposite. Their experiment can conclude that so-called blame-shifting
works but not that this is because of blame.

Modes offer a uniform way to compare the different semantics and  isolate
blame from the effect of the semantics' run-time checks. Specifically, the Natural and
Transient rational programmer come with a blame mode and an exception
mode. If the former succeeds debugging a program while the latter
fails, then it is safe to conclude that blame is indeed necessary for the
rational programmer. Put differently, the exception mode serves as the
{\em baseline\/} for blame's value within a given semantics; if the
programmer in this mode performs as well or better than the  blame one, a blame assignment
mechanism might be useless.
